Palomar 5 Gaps

### 2024-10-04 paper/main.tex
making colors of perturbers consistent

    It is also expected that a recent flyby would appear in most simulations, as uncertainties in a cluster's precise location increase as we look further back in time. In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions. The characteristic time scale is proportional to its orbital period. However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.

@@

    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation. However, 44 of these gaps are associated with NGC~2808. As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter. We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events. After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations. By integrating the binomial distribution, we compute that the probability of having at least 8 successes or fewer given 29 trials is approximately 1.2\%. While this outcome is unlikely it is not impossible. Additionally, by studying Fig.~\ref{fig:histogram_impact_time}, we notice that, although in much smaller numbers compared to NGC~2808, other pertubers cause gaps at the same time accross different simulations. These could also be violating the assumption of independent events.

@@

    \begin{figure}
      \centering
      \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
      \includegraphics[width=1\linewidth]{mass_size_plane.png}
      \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
      \caption{Characterisitcs of clusters that cause gaps compared to those that do not. \textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations. Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points. The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray. \textbf{Middle:} The mass-size plane. Uncertainties on the mass are indicated, while no uncertainties on the radius were reported. \textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}. The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
      \label{fig:mass_size_plane}
      \end{figure}

>>>>

    It is also expected that a recent flyby would appear in most simulations, as uncertainties in a cluster's precise location increase as we look further back in time. In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions. The characteristic phase-mixing time scale is proportional to its orbital period. However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.

@@

    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation. However, 44 of these gaps are associated with NGC~2808. As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter. We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events. After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations. By integrating the binomial distribution, we compute that the probability of having 8 successes or fewer given 29 trials is approximately 1.2\%. While this outcome is unlikely it is not impossible. Additionally, by studying Fig.~\ref{fig:histogram_impact_time}, we notice that, although in much smaller numbers compared to NGC~2808, other pertubers cause gaps at the same time accross different simulations. These could also be violating the assumption of independent events.

@@

    \begin{figure*}
      \centering
      \includegraphics[width=0.45\linewidth]{E_Lz_perturbers.png}
      \includegraphics[width=0.45\linewidth]{mass_size_plane.png}

      \includegraphics[width=0.45\linewidth]{impact_geometry_statistics_deltaP.png}
      \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
      \caption{Characterisitcs of clusters that cause gaps compared to those that do not. \textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations. Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points. The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray. \textbf{Middle:} The mass-size plane. Uncertainties on the mass are indicated, while no uncertainties on the radius were reported. \textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}. The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
      \label{fig:mass_size_plane}
      \end{figure*}
